A Gentle Introduction into Structural Causal Models

### 2021-05-07 topic_1_firstdraft.Rmd
restructure

To summarize, existing literature has provided an excellent introduction to structural causal models and bayesian causal networks. 
Judea Pearl (2009) provided a comprehensive and accessible introduction into this topic with his book on causality.
His work has addressed various misconceptions in social sciences, economics, causality and statistics. 
Peters et al. (2017) added to this discourse by addressing the relationship of causal modelling to physical sciences and causal inference for observational data.
Additionally, other scholars e.g. [@bareinboim2020pearl] have worked on the application of causal methods in machine learning. 
@tarka2018overview discuss the history of SEMs and SCMs.
This paper brings together these contributions. 
The objective of this paper is to provide a comprehensive summary of structural causal models and respective applications in social sciences, physical sciences and machine learning.
The insights in this paper are an abstraction of these pieces of work, offering the reading a basic understanding of the topic for future research and an overview of the current research.
The rest of the paper is structured as follows: 
The upcoming section studies the hierarchy of causation, looking at the different queries in the SCM and the BCN and what implications the functional/conditional probability specification carry.
The third section provides further reference to the intersection of machine learning and structural causal models. 
The fourth section studies historical differences and similarities between SCMs and causal models (SEM,PO) that emerged in the social sciences. 
The fifth section examines the relationship of physical sciences and SCMs, addressing the issue of time and modelling based on differential equations.

# The Causal Hierachy Compared

@@

# Causal Modelling and Machine Learning

# Causal Models and Social Sciences

Thereafter the Norwegian Economist Trygve Haavelmo (1944) provided a modelling framework for the field of economics in is paper 'The Probability Approach in Econometrics.'
One could argue that this paper provided a new of level of sophistication to the discipline of Econometrics as a distinct field. 
Haavelmo was awarded the Nobel Memorial Prize in Economic Science in 1989 for his work on this issue. 
His work focused on

# Causal Models and Physical Sciences

>>>>

To summarize, existing literature has provided an excellent introduction to structural causal models and bayesian causal networks. 
Judea Pearl (2009) provided a comprehensive and accessible introduction into this topic with his book on causality.
His work has addressed various misconceptions in social sciences, economics, causality and statistics. 
Peters et al. (2017) added to this discourse by addressing the relationship of causal modelling to physical sciences and causal inference for observational data.
Additionally, other scholars e.g. [@bareinboim2020pearl] have worked on the application of causal methods in machine learning. 
@tarka2018overview discuss the history of SEMs and SCMs.
This paper brings together these contributions. 
The objective of this paper is to provide a comprehensive summary of structural causal models and respective applications in social sciences, physical sciences and machine learning.
The insights in this paper are an abstraction of these pieces of work, offering the reading a basic understanding of the topic for future research and an overview of the current research.
The rest of the paper is structured as follows:

# Foundations of Structural Causal Models

Thereafter the Norwegian Economist Trygve Haavelmo (1944) provided a modelling framework for the field of economics in is paper 'The Probability Approach in Econometrics.'
One could argue that this paper provided a new of level of sophistication to the discipline of Econometrics as a distinct field. 
Haavelmo was awarded the Nobel Memorial Prize in Economic Science in 1989 for his work on this issue.

## Assumptions

Independence

## Pearl Causal Hierachy

@@

## Causal Inference and Time

## Graphical Models
